\textbf{Department Invited Speakers Do Not Reflect Trainee Diversity}

### 2019-09-13 submission/manuscript.Rmd
change methods and results to reflect change in race/ethnicity groupings

---
title: '**Department Invited Speakers Do Not Reflect Trainee Diversity**'
csl: mbio.csl
fontsize: 10pt
output:
  word_document: default
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE, include=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("knitr","rmarkdown", "markdown");
for (dep in deps){
  library(dep, verbose=FALSE, character.only=TRUE)
}
######################################################################
#-------------- Define the chunk options ----------------#
######################################################################
opts_chunk$set("tidy" = TRUE)
```

```{r LoadData, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
source("../code/load_data.R")
source("../code/get_figures.R")
```

@@

Supplemental:

@@

We hand-coded demographics using personal knowledge, photos, and CVs. The presenting gender of each individual was assigned using a binary system (man/woman). Diversity definitions vary according to the goals and population in question. However, in the United States, there is an inclination to consider both  together. We believe that it is important to distinguish between individuals of historically under-represented minority (HURM) and international backgrounds since each face different issues in the US and the academy and thus require different support systems. For instance, international scientists must contend with visa issues while HURMs have the trauma associated with living in a country who systematically shuts them out (despite an infrastructure that was built on their historical land and labor). For this reason, other assigned demographics included Caucasian, Historically Under-represented Minority (HURM), and International, each with a binary (yes/no) possibilty. Caucasian was assigned using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. HURM individuals were restricted to those with African-American, Indigenous and/or Hispanic heritage while International individuals were either visiting the US at the time of their seminar, or immigrated to the US as an adult.

## Results

To understand the representation of women, we compared the proportion of women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to `r gender_combined[[1,5]]`% of faculty hosts and `r gender_combined[[3,5]]`% of the invited speakers (Fig. 1A). Of 27 lectureships over the five year period, `r get_percent(10, 27)`% were awarded to women. The proportion of women as faculty hosts and speakers is equivalent to global estimates that 40% of microbiologists are women (Elseiver), with a slightly lower representation of women in lectureships.

Our analysis identified an over-representation of Caucasian individuals as hosting faculty and invited speakers, relative to the proportion of Caucasian trainees (Fig. 1B). We also observed declines in the representation of HURM and international faculty and speakers relative to the trainees, particularly postdocs (Fig 1B). Caucasian scientists also dominated lectureships, comprising `r get_percent(22, 27)`% of those awarded (Fig. 1C). Three and six lectureships were awarded to HURM and International scientists, respectively. Because the intersection of identities can compound biases and outcomes, we further examined the more prestigous lectureships by gender and Caucasian status. Caucasian men and women accounted for `r get_percent(12, 27)`% and `r get_percent(10, 27)`% of the lectureships, respectively, compared to `r get_percent(5, 27)`% non-Caucasian men and zero non-Caucasian women (Fig. 1D).

@@

Our data support the advancement of social role theory, which states that to improve retention of white women & HURMs, each group needs equivalent representation to counteract biases and improve self-efficacy. Implict biases that affect perceptions of marginalized groups are the primary issue, but we must acknowledge that it is not always possible to identify members of historically under-served communities. For instance, after data analysis, we learned that at least one speaker in our data set is a member of a HURM group, but it wasn't readily apparent from their internet presence or CV. This limitation makes two important points: that perspective is often as, if not more, important than self-identification with regards to biased outcomes, and that we need better tools to identify members of marginalized groups.

@@

We thank Dr. Harry Mobley and the Department of Microbiology & Immunology, University of Michigan for their input and financial support that enabled publication of our manuscript. We would also like to acknowledge Nick Lesniak and Dr. Ariangela Kozick for their comments and suggestions.

@@

## Code and data availability
The anonymized data, code for all analysis steps, and an Rmarkdown version of this manuscript is available at https://github.com/akhagan/Hagan_Libertucci_SpeakerDiversity_XXXX_2019/. Template and complete instructions for generating a field-specific Diversity website are available at https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/.

![Figure 1.](Figure_1.png)

>>>>

---
title: '**Department Invited Speakers Do Not Reflect Trainee Diversity**'
csl: mbio.csl
fontsize: 10pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
geometry: margin=1.0in
---

```{r setup, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE, include=FALSE}
#----------------- Read in necessary libraries -------------------#
deps = c("knitr","rmarkdown", "markdown");
for (dep in deps){
  library(dep, verbose=FALSE, character.only=TRUE)
}
#-------------- Define the chunk options ----------------#
opts_chunk$set("tidy" = TRUE)

#-------------- Read in data and figure ----------------#
source("../code/load_data.R")
source("../code/get_figures.R")
```

@@

@@

This is a retrospective study, so individuals were not asked for their identities. Instead we hand-coded proxy demographics using personal knowledge, photos, and CVs. The presenting gender of each individual was assigned using a binary system (man/woman). Due to the low number of individuals in the study, race/ethnicity demographics were split in three groups: Caucasian, Historically Under-represented Minority (HURM), and Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no) possibilty. Caucasian was assigned using the current U.S. Census definition where those of Middle Eastern, European, and Russian descent are included. HURM individuals were restricted to those with African-American, Indigenous, Alaskan/Hawaian Native, Latinx and/or Hispanic heritage. All others were placed into the NCNH group. We recognize that our proxy demographics are a limitation of the analysis and want to acknowledge that the presenting gender may not reflect the gender(s) that an individual identifies as. We also want to acknowledge that there are many other identities that are not captured in this limited analysis.

## Results
```{r result, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE, include=FALSE}
lect_summary <- lectureship %>% filter(value == "y") %>% 
  group_by(Speaker_Gender, demographic) %>% summarise(n = n()) %>% 
  mutate(prop = get_percent(n, sum(.$n)))

cauc_lect <- lect_summary %>% 
  filter(demographic == "Caucasian") %>% 
  summarise(sum = sum(n))
```
To understand the representation of women, we compared the proportion of women in each academic role. At the trainee level, more than half of students and postdoctoral fellows were women. That dropped to `r gender_combined[[1,4]]`% of faculty hosts and `r gender_combined[[3,4]]`% of the invited speakers (Fig. 1A). Of `r sum(lect_summary$n)` lectureships over the five year period, `r lect_summary[4,4]`% were awarded to women. The proportion of women as faculty hosts and speakers is equivalent to global estimates that 40% of microbiologists are women (Elseiver), with a slightly lower representation of women in lectureships.

Our analysis identified an over-representation of Caucasian individuals as hosting faculty and invited speakers, relative to the proportion of Caucasian trainees (Fig. 1B). We also observed declines in the representation of HURM and NCNH faculty and speakers relative to the trainees (Fig 1B). Caucasian scientists also dominated the more prestigous invited speaker lectureships, comprising `r get_percent(sum(cauc_lect$sum), sum(lect_summary$n))`% of those awarded (Fig. 1C). HURM and NCNH scientists were awarded `r lect_summary[[2,3]]` and `r lect_summary[[3,3]]` lectureships, respectively. Because the intersection of identities can compound biases and outcomes, we further examined the lectureships by gender and race/ethnicty status. Caucasian men and women accounted for `r lect_summary[[1,4]]`% and `r lect_summary[[4,4]]`% of the lectureships, respectively. Just `r 100 - get_percent(sum(cauc_lect$sum), sum(lect_summary$n))`% of lectureships were held by non-Caucasian men while none were held by non-Caucasian women (Fig. 1D).

@@

According to social role theory, this department is not adequietly serving their non-Caucasian trainees. Implict biases that affect perceptions of marginalized groups are the primary issue, but we must acknowledge that it is not always possible to identify members of historically under-served communities. For instance, after data analysis, we learned that at least one speaker in our data set should have been categorized as a HURM instead of Caucasian, but it wasn't readily apparent from their internet presence or CV. This limitation makes two important points: that percieved identity often plays a larger role than self-identification, and that we need better tools to identify members of marginalized groups.

@@

We thank Drs. Beth Moore and Harry Mobley and the Department of Microbiology & Immunology, University of Michigan for their input and financial support that enabled publication of our manuscript. We would also like to acknowledge Nick Lesniak and Dr. Ariangela Kozick for their comments and suggestions.

@@

## Code and data availability
The anonymized data, code for all analysis steps, and an Rmarkdown version of this manuscript is available at https://github.com/akhagan/Hagan_SpeakerDiversity_XXXX_2019/. Template and complete instructions for generating a field-specific Diversity website are available at https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/.

![**The demographics of invited speakers, hosting faculty, and trainees.** A) The proportion of women in each academic role. B) The proportion of each academic role represented by individuals that are Caucasian (left), Historically Underrepresented Minorities (HURM, center) or International (right). C-D)The percent of lectureships awarded to individuals that are C) Caucasian, HURM, International or D) Caucasian or Non-Caucasian by gender.](Figure_1.png)

### 2019-09-13 submission/manuscript.tex
change methods and results to reflect change in race/ethnicity groupings

We hand-coded demographics using personal knowledge, photos, and CVs.
The presenting gender of each individual was assigned using a binary
system (man/woman). Diversity definitions vary according to the goals
and population in question. However, in the United States, there is an
inclination to consider both together. We believe that it is important
to distinguish between individuals of historically under-represented
minority (HURM) and international backgrounds since each face different
issues in the US and the academy and thus require different support
systems. For instance, international scientists must contend with visa
issues while HURMs have the trauma associated with living in a country
who systematically shuts them out (despite an infrastructure that was
built on their historical land and labor). For this reason, other
assigned demographics included Caucasian, Historically Under-represented
Minority (HURM), and International, each with a binary (yes/no)
possibilty. Caucasian was assigned using the current U.S. Census
definition where those of Middle Eastern, European, and Russian descent
are included. HURM individuals were restricted to those with
African-American, Indigenous and/or Hispanic heritage while
International individuals were either visiting the US at the time of
their seminar, or immigrated to the US as an adult.

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers, relative to the proportion of
Caucasian trainees (Fig. 1B). We also observed declines in the
representation of HURM and international faculty and speakers relative
to the trainees, particularly postdocs (Fig 1B). Caucasian scientists
also dominated lectureships, comprising 81.48\% of those awarded (Fig.
1C). Three and six lectureships were awarded to HURM and International
scientists, respectively. Because the intersection of identities can
compound biases and outcomes, we further examined the more prestigous
lectureships by gender and Caucasian status. Caucasian men and women
accounted for 44.44\% and 37.04\% of the lectureships, respectively,
compared to 18.52\% non-Caucasian men and zero non-Caucasian women (Fig.
1D).

@@

Our data support the advancement of social role theory, which states
that to improve retention of white women \& HURMs, each group needs
equivalent representation to counteract biases and improve
self-efficacy. Implict biases that affect perceptions of marginalized
groups are the primary issue, but we must acknowledge that it is not
always possible to identify members of historically under-served
communities. For instance, after data analysis, we learned that at least
one speaker in our data set is a member of a HURM group, but it wasn't
readily apparent from their internet presence or CV. This limitation
makes two important points: that perspective is often as, if not more,
important than self-identification with regards to biased outcomes, and
that we need better tools to identify members of marginalized groups.

@@

We thank Dr.~Harry Mobley and the Department of Microbiology \&
Immunology, University of Michigan for their input and financial support
that enabled publication of our manuscript. We would also like to
acknowledge Nick Lesniak and Dr.~Ariangela Kozick for their comments and
suggestions.

@@

The anonymized data, code for all analysis steps, and an Rmarkdown
version of this manuscript is available at
\url{https://github.com/akhagan/Hagan_Libertucci_SpeakerDiversity_XXXX_2019/}.
Template and complete instructions for generating a field-specific
Diversity website are available at
\url{https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/}.

\begin{figure}
\centering
\includegraphics{Figure_1.png}
\caption{Figure 1.}
\end{figure}

>>>>

This is a retrospective study, so individuals were not asked for their
identities. Instead we hand-coded proxy demographics using personal
knowledge, photos, and CVs. The presenting gender of each individual was
assigned using a binary system (man/woman). Due to the low number of
individuals in the study, race/ethnicity demographics were split in
three groups: Caucasian, Historically Under-represented Minority (HURM),
and Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no)
possibilty. Caucasian was assigned using the current U.S. Census
definition where those of Middle Eastern, European, and Russian descent
are included. HURM individuals were restricted to those with
African-American, Indigenous, Alaskan/Hawaian Native, Latinx and/or
Hispanic heritage. All others were placed into the NCNH group. We
recognize that our proxy demographics are a limitation of the analysis
and want to acknowledge that the presenting gender may not reflect the
gender(s) that an individual identifies as. We also want to acknowledge
that there are many other identities that are not captured in this
limited analysis.

@@

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers, relative to the proportion of
Caucasian trainees (Fig. 1B). We also observed declines in the
representation of HURM and NCNH faculty and speakers relative to the
trainees (Fig 1B). Caucasian scientists also dominated the more
prestigous invited speaker lectureships, comprising 81.48\% of those
awarded (Fig. 1C). HURM and NCNH scientists were awarded 3 and 2
lectureships, respectively. Because the intersection of identities can
compound biases and outcomes, we further examined the lectureships by
gender and race/ethnicty status. Caucasian men and women accounted for
44.44\% and 37.04\% of the lectureships, respectively. Just 18.52\% of
lectureships were held by non-Caucasian men while none were held by
non-Caucasian women (Fig. 1D).

@@

According to social role theory, this department is not adequietly
serving their non-Caucasian trainees. Implict biases that affect
perceptions of marginalized groups are the primary issue, but we must
acknowledge that it is not always possible to identify members of
historically under-served communities. For instance, after data
analysis, we learned that at least one speaker in our data set should
have been categorized as a HURM instead of Caucasian, but it wasn't
readily apparent from their internet presence or CV. This limitation
makes two important points: that percieved identity often plays a larger
role than self-identification, and that we need better tools to identify
members of marginalized groups.

@@

We thank Drs. Beth Moore and Harry Mobley and the Department of
Microbiology \& Immunology, University of Michigan for their input and
financial support that enabled publication of our manuscript. We would
also like to acknowledge Nick Lesniak and Dr.~Ariangela Kozick for their
comments and suggestions.

@@

The anonymized data, code for all analysis steps, and an Rmarkdown
version of this manuscript is available at
\url{https://github.com/akhagan/Hagan_SpeakerDiversity_XXXX_2019/}.
Template and complete instructions for generating a field-specific
Diversity website are available at
\url{https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/}.

\begin{figure}
\centering
\includegraphics{Figure_1.png}
\caption{\textbf{The demographics of invited speakers, hosting faculty,
and trainees.} A) The proportion of women in each academic role. B) The
proportion of each academic role represented by individuals that are
Caucasian (left), Historically Underrepresented Minorities (HURM,
center) or International (right). C-D)The percent of lectureships
awarded to individuals that are C) Caucasian, HURM, International or D)
Caucasian or Non-Caucasian by gender.}
\end{figure}
